---
title: "R Data Wrangling Exercise - Kelly Hatfield"
output:
  html_document:
    toc: FALSE
---

# MUGSI Data


We accessed this data from [CDC HAIC Data Viz](https://www.cdc.gov/hai/eip/haicviz.html)

### Multi-site Gram-negative Surveillance Initiative

The healthcare-associated infection component of CDC's Emerging Infections Program engages a network of state health departments and their academic medical center partners to help answer critical questions about emerging threats, advanced infection tracking methods, and antibiotic resistance in the United States. Information gathered through this activity will play a key role in shaping future policies and recommendations targeting HAI prevention.

 Selected gram-negative bacteria are under surveillance, as they are becoming resistant to all or nearly all antibiotics, meaning that patients with infections from these bacteria might have few or no treatment options. Infections due to highly resistant bacteria, such as carbapenem-resistant Enterobacterales (CRE), carbapenem-resistant *Acinetobacter baumannii* (CRAB)are mainly associated with healthcare settings and have high death rates.

We will explore longitudinal data from these pathogens on CDC's website.

## R Code for loading packages

```{r}

#Load Tidyverse
library(tidyverse) 
library(ggplot2)

mugsi <- read_csv("Data_Exercise/Data/MUGSI.csv")

#view some data
print(mugsi)

#get an overview of data structure
str(mugsi)
#get a summary of data
summary(mugsi)



#look at values for organism, topic, viewby, and series and 
table(mugsi$Organism)
table(mugsi$Topic)


```

## Analytic Goals

The objective of this analysis will be to view the number of CRE cases by year and patient location. We will first subset the data to only include that information.

However, the data are laid out in a funny shape for analysis. For this project we will subset the data to be laid out in a more Tidy manner.

We will make columns year, organism, case rates, and death rates. There will be a row for each year, organism, and location combination.

```{r}

mugsi_analysis0 = mugsi%>% filter(Topic=="Case Rates") %>% filter(Viewby=="Patient location") %>%rename(patient_location = Series) %>%  rename(case_rates = Value)

mugsi_analysis0a <- mugsi_analysis0[,!names(mugsi_analysis0) %in% c("Topic", "Viewby")]

mugsi_analysis1 = mugsi%>% filter(Topic=="Death Rates") %>%rename(patient_location = Series) %>%  rename(death_rates = Value)

mugsi_analysis1a <- mugsi_analysis1[,!names(mugsi_analysis1) %in% c("Topic", "Viewby")]
  
mugsi_tidy <- left_join(mugsi_analysis0a, mugsi_analysis1a, by=c("YearName", "Organism", "patient_location"))

print(mugsi_tidy)

view(mugsi_tidy)
```

## Analysis Next Steps

In order to plot the number of CRE cases by patient location over time you will need the following variables:

YearName : Numeric variable for year Organism: Indicates if CRE or CRAB data case_rates: Numeric variable describing organism rates per 100,000 population patient_location: categorical variable describing patient location (LTAC= long term acute care hospital, and LTCF= long term care facility, All Cases = Sum of all locations

### Analysis notes:

You should look at either "All cases" or the four location groupings (hospital inpatient, community, LTCF, and LTAC). "All cases" represents the sum of the four subset locations.

Death rates are only available for "All cases" (not stratified by patient location).

### Kelly's Analysis

I am going to graph CRE rates over time by location.

```{r}

mugsi_plot1 <- subset(mugsi_tidy, Organism=="CRE")

ggplot(mugsi_plot1, aes(x=YearName, y=case_rates, color=patient_location, group=patient_location)) + geom_point() + geom_line() +labs(y="Case Rates per 100,000 population", x="Year", title = "Annual CRE Rates by Patient Location")
                         



```

Now I will save this into a R dataset for you to analyze!

```{r}

#Export

save(mugsi_tidy, file = "Data_Exercise/Data/MUGSI_tidy.RData") #raw and clean data if needed
```

### You are up!

Some potential analytic ideas:

Look at trends in case rates and death rates for each year for each pathogen group.

Look to see if CRAB trends are similar to CRE trends in specific settings.
