---
title: "Tidy Tuesday Exercise 2"
output: 
  html_document:
    toc: FALSE
---

# Tidy Tuesday Part 2: Week of April 11, 2023

This weeks data is called: [US Egg Production Data 2007-2021](https://osf.io/z2gxn/) and comes from [The Humane League Labs US Egg Production Dataset](https://thehumaneleague.org/article/E008R01-us-egg-production-data).

## Accessing Data

```{r}
# Get the Data

# Read in directly

eggproduction  <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-04-11/egg-production.csv')
cagefreepercentages <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-04-11/cage-free-percentages.csv')


#get a summary of data
summary(eggproduction)
summary(cagefreepercentages)


#Add some libraries

library(ggplot2)
library(tidyverse)

```

## Data Dictionary

I am copying and pasting the data dictionaries from the Tidy Tuesday github for reference here:

### `egg-production.csv`

| variable       | class     | description                                                                                                                                             |
|----------------|-----------|---------------------------------------------------------------------------------------------------------------------------------------------------------|
| observed_month | double    | Month in which report observations are collected,Dates are recorded in ISO 8601 format YYYY-MM-DD                                                       |
| prod_type      | character | type of egg product: hatching, table eggs                                                                                                               |
| prod_process   | character | type of production process and housing: cage-free (organic), cage-free (non-organic), all. The value 'all' includes cage-free and conventional housing. |
| n_hens         | double    | number of hens produced by hens for a given month-type-process combo                                                                                    |
| n_eggs         | double    | number of eggs producing eggs for a given month-type-process combo                                                                                      |
| source         | character | Original USDA report from which data are sourced. Values correspond to titles of PDF reports. Date of report is included in title.                      |

### `cage-free-percentages.csv`

| variable       | class     | description                                                                                                                                           |
|----------------|-----------|-------------------------------------------------------------------------------------------------------------------------------------------------------|
| observed_month | double    | Month in which report observations are collected,Dates are recorded in ISO 8601 format YYYY-MM-DD                                                     |
| percent_hens   | double    | observed or computed percentage of cage-free hens relative to all table-egg-laying hens                                                               |
| percent_eggs   | double    | computed percentage of cage-free eggs relative to all table eggs,This variable is not available for data sourced from the Egg Markets Overview report |
| source         | character | Original USDA report from which data are sourced. Values correspond to titles of PDF reports. Date of report is included in title.                    |

### 

\
#Some cleaning

```{r}

#Separating out hatching and table eggs

hatching_eggsnhens <- subset(eggproduction, prod_type=='hatching eggs',c(observed_month, prod_process, n_eggs, n_hens)) 
hatching_eggsnhens$eggsperhen <- hatching_eggsnhens$n_eggs / hatching_eggsnhens$n_hens

hatching_eggsnhens

table_eggsnhens <- subset(eggproduction, prod_type=='table eggs',c(observed_month,prod_process, n_eggs, n_hens)) 
table_eggsnhens$eggsperhen <- table_eggsnhens$n_eggs / table_eggsnhens$n_hens

tablehens1<- subset(eggproduction, prod_type=='table eggs',c(observed_month, prod_process, n_hens))  
tablehens2 <- tablehens1 %>% pivot_wider(names_from = prod_process, values_from = n_hens)%>% rename("all-table-n_hens" = "all", "cage-free-organic-table-n_hens" = "cage-free (organic)",  "cage-free-non-organic-table-n_hens" = "cage-free (non-organic)"       )


tableeggs1 <- subset(eggproduction, prod_type=='table eggs',c(observed_month, prod_process, n_eggs))  
tableeggs2 <- tableeggs1 %>% pivot_wider(names_from = prod_process, values_from = n_eggs)%>% rename("n_eggs_all" = "all", "n_eggs_cfo" = "cage-free (organic)",  "n_eggs_cfno" = "cage-free (non-organic)")



combined_data <-full_join(hatching_eggsnhens, table_eggsnhens, by = "observed_month")



cleandata <-full_join(combined_data, cagefreepercentages, by = "observed_month") %>% rename("percent all table egg laying hens cage free" = "percent_hens", "percent all table eggs organic" = "percent_eggs")                                                               
```

## Data Explorations

#### Lets look at table hatching eggs and hens over time

```{r}

ggplot(hatching_eggsnhens, aes(x=observed_month, y=eggsperhen, color="pink")) +geom_path() +geom_point()
       

```

It appears certain months have low egg production

```{r}
#format as month
hatching_eggsnhens$month <- format(as.Date(hatching_eggsnhens$observed_month, format="%d/%m/%Y"),"%m")

ggplot(hatching_eggsnhens, aes(x =month , y = eggsperhen)) + geom_boxplot(fill = "grey92") + geom_point(size = 2, alpha = .15,position = position_jitter(seed = 1, width = .2)) 

```

Looks like something is weird in February: the median \# of eggs per month is closer to 17, compared to \>18 in all other months. Maybe this is due to an increased desire for more table eggs in those months.

Maybe we'll use this in our model. But lets also explore some table egg information.

#### Lets look at the number of table eggs per hen over time

```{r}

table_eggsnhensall <-subset(table_eggsnhens,prod_process=='all')

ggplot(table_eggsnhensall, aes(x=observed_month, y=eggsperhen, color="pink")) +geom_path() +geom_point()
       
#format as month
table_eggsnhensall$month <- format(as.Date(table_eggsnhensall$observed_month, format="%d/%m/%Y"),"%m")

ggplot(table_eggsnhensall, aes(x =month , y = eggsperhen)) + geom_boxplot(fill = "grey92") + geom_point(size = 2, alpha = .15,position = position_jitter(seed = 1, width = .2)) 
```

THe number of eggs per hen for table eggs is also lower in February! Strange. Its also interesting to note that the number of eggs per hen is much higher for table eggs than hatching eggs!

#### Lets also look at the number of table eggs in total over time

```{r}

ggplot(tableeggs1, aes(x=observed_month, y=n_eggs, 
                      group=prod_process, color=prod_process)) +geom_path() +geom_point()
       
       
```

Some key take aways:

1.  Interestingly, looks like all table eggs was on an upward trend... until 2020.

2.  Cage free non-organic eggs seem to be increasing production faster than organic eggs

Some potential analyses:

1.  Potential analysis: ITS model looking at linear trend in egg production until March 2020 (start of COVID-19 pandemic)

2.  Difference in Difference model looking at slope of organic versus non-organic eggs

Lets look at a pre COVID-19 and post-COVID-19

```{r}


#Add covid indicator for any month in March 2020 forward

tableeggs2$cdate <- as.Date("2020-02-28")
tableeggs2$covid_ind = ifelse(tableeggs2$observed_month>tableeggs2$cdate,1,0)
tableeggs2$n_eggs_all_m <- tableeggs2$n_eggs_all/1000000


tableeggs2$mindate = min(tableeggs2$observed_month)
tableeggs2$daysmodel = as.numeric (tableeggs2$observed_month-tableeggs2$mindate)
tableeggs2$interaction <- ifelse(tableeggs2$observed_month > tableeggs2$cdate,tableeggs2$daysmodel,0)

#Plotting interrupted time series model
ggplot(tableeggs2, aes(x=observed_month, y=n_eggs_all_m, 
                      group=factor(covid_ind),
                      color=factor(covid_ind))) +geom_point() + geom_vline(xintercept = tableeggs2$cdate, linetype="dashed", 
                color = "gray2", linewidth=1)+ geom_smooth(method="lm", se=FALSE) + ggtitle("Plot of table eggs produced over time") + xlab("Date") + ylab("Number of eggs (in millions)")

```

## Model approach

#### Question: Was there a change in egg production related to the COVID-19 pandemic?

We are going to take an interrupted time series approach to understand if the linear trend in egg production changes beginning in March 2020, at the start of COVID-19 pandemic.

#### Here are my steps: 

1.  Data cleaning for the model: See above.

2.  Assign 70% of data to training and 30% to test (make assignments by COVID_ind status so we have the same amount of data pre-post intervention)

3.  Calculate linear regression with the following covariates: observed_month(month in time series), covid_ind (a 0/1 indicator variable for if in teh COVID-19 time period), and an interaction between covid_ind and observed_mont

4.  Interpretation will be as follows:

    1.  Parameter (observed_month): Slope (Trend) before the pandemic

    2.  Parameter estimate (COVID_IND): Immediate change in egg production at start of pandemic

    3.  Parameter estimate (interaction term): Change in slope (trend) during the pandemic

5.  *Repeat model on test data to see if interpretation of parameter estimates are similar*

6.  **Summarize findings**

```{r}

#Modeling Approach
library(tidyverse)
library(tidymodels)
library(performance)

set.seed(777)

data_split <- initial_split(tableeggs2, prop = 2/3, strata=covid_ind)

# Create data frames for the two sets:
train_data <- training(data_split)
#36 obs
test_data  <- testing(data_split)
#20 obs

model_recipe <- recipe(n_eggs_all_m ~ daysmodel + covid_ind + interaction, data= train_data)

#model: linear regression using GLM engine
ln_mod <- linear_reg() %>% set_engine ("lm")
wflow <- workflow() %>% add_model(ln_mod) %>% add_recipe(model_recipe)

#Fitting the model to train dataset
model_fit_train <-  wflow %>% fit(data = train_data)
#Looking at model output;
model_fit_train %>% 
  extract_fit_parsnip() %>% 
  tidy()



#Applying the model to unseen test data
predict(model_fit_train, test_data)

#augment data
aug <- augment(model_fit_train, test_data)

#pull month and egg prediction
observed_daysmodel <- aug %>% pull(daysmodel)
pred <- aug %>% pull(.pred)

ggplot(aug, aes(x=.pred, y= n_eggs_all_m, group=factor(covid_ind), color=factor(covid_ind))) +
  geom_point() +
  geom_abline(intercept=0, slope=1) +
  labs(x='Predicted Values', y='Actual Values', title='Predicted vs. Actual Values')
#run rmse
rmse_vec(observed_daysmodel, pred)

```

#### Summary of findings:

From our training dataset we estimate an increase of 0.74 million eggs per day prior to COVID-19. At the start of COVID-19 we saw an immediate decrease of 650 million eggs per day! Then, after COVID-19 we saw an increased trend of 0.18 million additional eggs per day (for a post-COVID-19 trend of 0.92 million more eggs produced each day!!).

#### Summary of our model fit: 

In our plot of observed versus predicted values for the test data we feel confident that our model predicted well. There was potentially less good prediction in teh COVID-19 era than the pre-COVID-19 era but that is likely due to the small sample of data in that period for the model.
