---
title: "exploration"
editor: visual
---

## Identify Variables

```{r}
library(here)
library(tidyverse)
library(tidymodels)

CleanSymp <- readRDS("~/Documents/MADA 2023/kellyhatfield-MADA-portfolio/fluanalysis/Data/CleanSymp.Rds")
ls(CleanSymp)
summary(CleanSymp)



```

Things to note:

-   Most variables are categorical Yes/No.

-   Temperature is continuous from 97.2 to 103.1

-   Weakness, CoughIntensity, and Myalgia are scored None, Mild, Moderate, Severe

-   MyalgiaYN, CoughYN2, and WeaknessYN are all Yes/No versions of their corresponding intensity variable (None, Mild, Moderate, Severe)

## Data Explorations: Body Temperature

First, we want to look at temperature with a few key variables. We have selected cough, chest pain and wheeze.

```{r}

CleanSymp %>% summarize(min=min(BodyTemp),mean=mean(BodyTemp), q1 = quantile(BodyTemp, 0.25), median = mean(BodyTemp),  q3 = quantile(BodyTemp, 0.75), max=max(BodyTemp))

CleanSymp %>% group_by(CoughYN2) %>% summarize(mean=mean(BodyTemp), q1 = quantile(BodyTemp, 0.25), median = mean(BodyTemp),  q3 = quantile(BodyTemp, 0.75))

ggplot(CleanSymp, aes(x = CoughYN2, y = BodyTemp)) + geom_boxplot(fill = "grey92") + geom_point(size = 2, alpha = .15,position = position_jitter(seed = 1, width = .2)) 

CleanSymp %>% group_by(ChestPain) %>% summarize(mean=mean(BodyTemp),  q1 = quantile(BodyTemp, 0.25), median = median(BodyTemp),  q3 = quantile(BodyTemp, 0.75))

ggplot(CleanSymp, aes(x = ChestPain, y = BodyTemp)) + geom_boxplot(fill = "grey92") + geom_point(size = 2, alpha = .15,position = position_jitter(seed = 1, width = .2)) 

CleanSymp %>% group_by(Wheeze) %>% summarize(mean=mean(BodyTemp),  q1 = quantile(BodyTemp, 0.25), median = median(BodyTemp),  q3 = quantile(BodyTemp, 0.75))
ggplot(CleanSymp, aes(x = Wheeze, y = BodyTemp)) + geom_boxplot(fill = "grey92") + geom_point(size = 2, alpha = .15,position = position_jitter(seed = 1, width = .2)) 

```

Since median teperature seems to be slightly elevated for the cough variable, we will look at it further for those varying rankings of the intensity of the cough.

```{r}

CleanSymp %>% group_by(CoughIntensity) %>% summarize(mean=mean(BodyTemp),  q1 = quantile(BodyTemp, 0.25), median = median(BodyTemp),  q3 = quantile(BodyTemp, 0.75))

ggplot(CleanSymp, aes(x = CoughIntensity, y = BodyTemp)) + geom_boxplot(fill = "grey92") + geom_point(size = 2, alpha = .15,position = position_jitter(seed = 1, width = .2)) 

```

The mean body temperature doesn't seem to vary too much by increasing cough intensity group. However, perhaps some increase in median or q3 values.

## Data Explorations: Nausea

For nausea we have decided to assess the relationship of nausea with subjective fever, myalgia, and abdominal pain.

```{r}
# Variables of interest with Nausea 
table1 <- table(CleanSymp$Nausea,CleanSymp$SubjectiveFever)
table1
prop.table(table1) %>% {.*100} %>% round(2)

table2 <- table(CleanSymp$Nausea,CleanSymp$MyalgiaYN)
table2
prop.table(table2) %>% {.*100} %>% round(2)
table3 <- table(CleanSymp$Nausea,CleanSymp$AbPain)
table3
prop.table(table3) %>% {.*100} %>% round(2)

```

## Data Explorations: Nausea and Body Temp

FFinally lets look at the relationships between our two primary variables.

```{r}

CleanSymp %>% group_by(Nausea) %>% summarize(mean=mean(BodyTemp),  q1 = quantile(BodyTemp, 0.25), median = median(BodyTemp),  q3 = quantile(BodyTemp, 0.75))

ggplot(CleanSymp, aes(x = Nausea, y = BodyTemp)) + geom_boxplot(fill = "grey92") + geom_point(size = 2, alpha = .15,position = position_jitter(seed = 1, width = .2)) 

```
