---
title: "wrangling"
author: "Kelly Hatfield"
format: html
editor: visual
---

## Step 1: Opening Data

Working to find the data

```{r}
library(here)
library (tidyverse)


raw_data <- readRDS(file = "/Users/kellymccormickhatfield/Documents/MADA 2023/kellyhatfield-MADA-portfolio/fluanalysis/Data/SympAct_Any_Pos.Rda")

```

## Step 2: Wrangling Data

Part 1: Subsetting Data

-   Remove all variables that have *Score* or *Total* or *FluA* or *FluB* or *Dxname* or *Activity* in their name. 

-   Also remove the variable `Unique.Visit`. You should be left with 32 variables coding for presence or absence of some symptom. Only one, temperature, is continuous. A few have multiple categories.

-   Remove any NA observations, there aren't many.

```{r}

#List variable names

ls(raw_data)

raw_data2 <- raw_data %>% select(-contains(c("Score","Total","FluA","FluB","Dxname", "Activity"))) %>% select(-contains(c("Unique.Visit")))

#Contains 32 variables. Yay!

#Remove missing observations
raw_data3 <- na.omit(raw_data2)
#Only dropped 5 observations. 

summary(raw_data3)


```

## Step 3: Manipulating Data

### Categorical/Ordinal predictors

```{r}

```

### Deleting Repetitive Variables

```{r}
summary(raw_data3)
 #Delete Variables with yes/no observations that  are represented with a different variable
Fludata1 <- select(raw_data3,-c(CoughYN, WeaknessYN, CoughYN2, MyalgiaYN))

#Drop variables <50 Y/N

Fludata2 <- select(Fludata1, -c(Hearing, Vision))

```

## Step 3: Save Wrangled Data as a RDS

```{r}
here()
#Need to add some files to here


path <- here("fluanalysis","Data","CleanSymp.Rds")
saveRDS(raw_data3, file = path)

path <- here("fluanalysis", "Data", "FinalDataML.Rds")
saveRDS(Fludata2, file = path)
```
